---
title: "EDS241: Assignment 1"
author: "Connor Flynn"
date: '`r format(Sys.time(), "%m/%d/%Y")`'
output: 
  pdf_document:
    toc: false
    number_sections: yes
header-includes:
  - \setlength{\parindent}{1em}
  - \usepackage{float}
--- 

``` {r setup, echo = FALSE, message = FALSE, warning = FALSE}

# set default chunk options
knitr::opts_chunk$set(fig.width = 4, fig.height = 3, 
                      echo = TRUE, message = FALSE, warning = FALSE)


# load packages
packages=c("stargazer", "here", "tidyr", "dplyr","stringr", "janitor", 
           "cowplot", "ggplot2", "tinytex", "datasets", "tibble", "openxlsx", 
           'estimatr', "knitr", "car")

for (i in packages) {
  if (require(i,character.only=TRUE)==FALSE) {
    install.packages(i,repos='http://cran.us.r-project.org')
  }
  else {
    require(i,character.only=TRUE)
  }
}

#devtools::install_github('rstudio/rmarkdown')
options(scipen=20) # not scientific notation
```


```{r}
data <- read.xlsx(here('data/CES4.xlsx'))
```


```{r}
data <- data %>% 
  select(Census.Tract, Total.Population, California.County, Low.Birth.Weight, PM2.5, Poverty)
```



```{r}
PM2.5_Concentration_California_Mean <- mean(data$PM2.5)
PM2.5_Concentration_California_Mean
```

A). The average concentration of PM2.5 acrosss all census tracts in California is 10.1527 micro grams /cubic meter.




```{r}
Poverty_California_County <- data %>% 
  group_by(California.County) %>% 
  summarise(Poverty_Mean_County = mean(Poverty, na.rm = TRUE))
```

B). Based on mean poverty at the county level, Tulare has the highest level of poverty in California.


C).
```{r}
Low.Birth.Weight_hist <- ggplot(data = data, aes(x = Low.Birth.Weight)) +
  geom_histogram()

Low.Birth.Weight_hist
```


```{r}
PM2.5_hist <- ggplot(data = data, aes(x = PM2.5)) +
  geom_histogram()

PM2.5_hist
```


```{r}
model_1 <- lm_robust(Low.Birth.Weight ~ PM2.5, data = data)
summary(model_1)
```

D). Estimated Slope Coefficient = 0.1179
    Heteroskedasticity-robust standard error = 0.008402
    For every one unit increase in PM2.5 concentration, there is a 0.1179 increase in percentage of low birth weights
    Based on the p value, the effect of PM2.5 on LowBirthWeight is statistically significant at the 95% Confidence level




```{r}
model_3 <- lm_robust(Low.Birth.Weight ~ PM2.5 + Poverty, data = data)
summary(model_3)
```

F). For every one unit increase in poverty, there is a 0.02 increase in percentage of low birth weights.
    The estimated coefficient on PM2.5 decreases compared to the regression excluding poverty as an explanatory variable (D).       
    This makes sense because we've added an explanatory variable that effects the percentage of low birth weights, therefore decreasing the                    significance of PM2.5 in the regression. 
    

```{r}
linearHypothesis(model_3,c("PM2.5 = Poverty"), white.adjust = "hc2")
```


G). Based on the linear hypothesis test above, we can reject the null hypothesis that the effect of PM2.5 is equal to the
effect of Poverty in regression F.























Exploratory Visualization & Analysis


```{r}
birth_weight_PM2.5_plot = ggplot(data = data, aes(x = PM2.5, 
                                                  y = Low.Birth.Weight)) +
  geom_point()+
  geom_smooth(method  = "lm", col = "blue")

birth_weight_PM2.5_plot
```



```{r}
model_5 <- lm_robust(Low.Birth.Weight ~ Poverty, data = data)
summary(model_5)
```



```{r}
birth_weight_Poverty_plot = ggplot(data = data, aes(x = Poverty, 
                                                  y = Low.Birth.Weight)) +
  geom_point()+
  geom_smooth(method  = "lm", col = "blue")

birth_weight_Poverty_plot
```

